### 2022-03-28 scripts/scratch.rmd
Reshuffle folders + README

```{r}
library(gssr)
library(tidyverse)
# don't recommend getting single year -- not enough data
# gss21 <- gss_get_yr(2021)
# gss21["sexsex"]

```

```{r}
data(gss_all)
```

```{r}
gss_clean <- gss_all |>
  mutate(
    id = paste0(year, '_', id),
    sexnow = coalesce(sexnow, sexnow1),
    sexbirth = coalesce(sexbirth, sexbirth1),
    )# |>
  #select(id, year, sex, sexnow, sexbirth, sexornt, sexsex, sexsex5, ptnrornt)
```
```{r}

```

```{r}
# grab a subset of data, specifically
# all responses from years in which all
# required questions in {qns} are asked/
# have answers.
# Optionally, also filter to only entries
# with answers to all questions.
getSubsetWithAnswers <- function(qns, clean_na = FALSE) {
  years_all_asked <- gss_all |>
    gss_which_years(qns) |>
    na_if(FALSE) |>
    drop_na() |>
    pull(year)

  gss_all |>
    filter(year %in% years_all_asked) |>
    select(year, id, all_of(qns)) |>
    filter(across(qns, ~ !clean_na | !is.na(.)))
}
```

```{r}
# this dataset contains a whopping 1 intersex person,
# and only 12 who identify their SEXNOW as 3 [transgender].
# In fact, only 42 people identify their SEXBIRTH as different from SEXNOW.
getSubsetWithAnswers(c("sexnow", "sexbirth"), clean_na = T) |>
  ggplot(aes(x = as.factor(sexbirth), y = as.factor(sexnow))) +
  geom_bin2d()
```
```{r}
# Interesting...
gss_clean |> 
  filter(sexbirth != sexnow |
         sexnow != sex |
         sex != sexbirth) |> 
  select(id, sexbirth, sexnow, sex)
```

```{r}
subsetWith(c('sexsex', 'sexnow', 'sexornt')) |>
  ggplot(aes(x = as_factor(sexsex), fill=as_factor(sexsex))) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('blue', 'purple', 'red')) +
  theme(axis.text.x = element_blank()) +
  facet_grid(as_factor(sexnow) ~ as_factor(sexornt),
             scales = "free_y",
             labeller = label_wrap_gen(width = 16, multi_line = T)
             )
```
```{r}
gss_clean |>
  filter(!if_any(c(sexsex5, sexnow, sexornt), is.na)) |>
  ggplot(aes(x = as_factor(sexsex5), fill=as_factor(sexsex5))) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('blue', 'purple', 'red')) +
  #geom_text(stat="count", aes(label=..count..), position = position_stack(vjust=.5), color="black") +
  theme_bw() + 
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank()
    ) +
  facet_grid(as_factor(sexornt) ~ as_factor(sexnow),
             scales = "free_y",
             labeller = label_wrap_gen(width = 16, multi_line = T)
             )
```

```{r}
# Well, let's see just /how/ flawed the data is.

subsetWith <- function(..) {
  gss_clean |>
    filter(!if_any(.., is.na))
}
```
```{r}
gender_counts <- subsetWith(c('sex', 'sexbirth', 'sexnow')) |>
  mutate(
    sex_changed = sexbirth != sexnow,
    sex_mismatch = sexbirth == sexnow & sexnow != sex,
    sex_coded_birth = sex == sexbirth,
    sex_coded_now = sex == sexnow
  ) |>
  count(sex, sexbirth, sexnow, sex_changed, sex_mismatch, sex_coded_birth, sex_coded_now) |>
  mutate(
  trans_miscoded = sex_changed & !sex_coded_now
  )

gender_counts |>
  ggplot(aes(x = as_factor(trans_miscoded), y = n, fill=trans_miscoded)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Respondent-identified gender",
    y = "Interviewer-coded sex of respondent",
    fill = "Apparent misgendering?"
  ) +
   theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  facet_grid(as_factor(sex) ~ as_factor(sexnow), scales="free_y", labeller = label_wrap_gen(width = 16, multi_line = T))

# gss_clean |> 
#   filter(!if_any(c(sex, sexbirth, sexnow), is.na)) |>
#   mutate(
#     sex_mismatch = 
#   )
#   group_by(year) |>
#   summarize()
#   ggplot(aes(x = factor(year)))
```
```{r}
gender_counts |>
  #filter(sexnow %in% c(1, 2)) |>
  ggplot(aes(x = sex_coded_now, y = n, fill = sex_coded_now)) +
  geom_bar(stat = "identity", position="dodge") +
  geom_text(aes(label=n, vjust="center")) +
  labs(
    x = "Assigned sex at birth (respondent-identified)",
    y = "Gender identity (respondent-identified)*",
    fill = "Reporter-indicated sex matches gender?"
  ) +
   theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  #scale_fill_manual(values=c('blue', 'red')) + 
  facet_grid(as_factor(sexnow) ~ as_factor(sexbirth), scales="free_y", labeller = label_wrap_gen(width = 16, multi_line = T))
```

```{r}
sex_codes <- c('M', 'F', 'I')
gender_codes <- c('M', 'F', 'T', 'X')

subsetWith(c('sex', 'sexbirth', 'sexnow')) |>
  mutate(
    code = paste0(sex_codes[sexbirth], gender_codes[sexnow])
    ) |>
  select(code, sex) |>
  group_by(code) |>
  table() |>
  knitr::kable(col.names = c("Coded male", "Coded female")) |>
  kableExtra::kable_classic()
```

```{r}
subsetWith(c('sexornt', 'ptnrornt')) |>
  select('sexornt', 'ptnrornt') |>
  filter(ptnrornt != 4) |> # 'I've never had a sexual partner'
  group_by(sexornt) |>
  ggplot(aes(x = as_factor(sexornt), y = as_factor(ptnrornt))) + 
  geom_bin_2d() +
  scale_fill_viridis_c(option="C")
```

```{r}
subsetWith(c('sexnow', 'sexornt', 'ptnrornt', 'ptnrsxnow')) |>
  select('sexnow', 'sexornt', 'ptnrornt', 'ptnrsxnow') |>
  ggplot(aes(x = as_factor(ptnrornt), fill = as_factor(ptnrsxnow))) +
  geom_histogram(stat='count') +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  labs(
    title = "Gender identity of subject",
    x = "Sexual orientation of last sex partner (Gay, Bisexual, Straight)",
    y = "Sexual orientation of subject",
    fill = "Gender identity of last sex partner"
  ) +
  facet_grid(as_factor(sexornt) ~ as_factor(sexnow), 
  scale="free_y",          
  labeller = label_wrap_gen(width = 16, multi_line = T))
```

```{r}
# based on `sex`, how many respondents have primarily been having
# sexual relationships with an unexpected cohort? (e.g., straight men primarily having sex with men)
subsetWith('sex', 'sexornt', 'sexsex') |>
  select(sex, sexornt, sexsex) |>
  mutate(
    expected_sexsex = case_when(
      sex == 1 & sexornt == 3 ~ 3,
      sex == 1 & sexornt == 1 ~ 1,
      sex == 2 & sexornt == 3 ~ 1,
      sex == 2 & sexornt == 1 ~ 3,
    )
  ) |>
  filter(
    #!is.na(expected_sexsex) &
    expected_sexsex != sexsex
  )
```
```{r}
# sim. with sexnow and sexsex5
subsetWith('sexnow', 'sexornt', 'sexsex5') |>
  select(year, sexnow, sexornt, sexsex5) |>
  rowwise() |>
  mutate(
    label = paste(
      c("Gay", "Bisexual", "Straight")[sexornt],
      c("men", "women", "transgender people", "people of other genders")[sexnow],
      "primarily having sex with",
      c("men", "women", "both men and women")[sexsex5]
    , collapse = " "),
    expected_sexsex = case_when(
      sexnow == 1 & sexornt == 3 ~ 3,
      sexnow == 1 & sexornt == 1 ~ 1,
      sexnow == 2 & sexornt == 3 ~ 1,
      sexnow == 2 & sexornt == 1 ~ 3,
    )
  ) |>
  count(label)

# |>
#   filter(
#     #!is.na(expected_sexsex) &
#     sexsex != expected_sexsex
#   )
```